---
title: "Diamond Exploration Tool Documentation"
author: "Chelsi Snellen"
date: "January 20, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This tool is designed to make the data exploration process for the diamonds dataset easier for people who know little to no coding. 

With this tool, you can choose three aspects of the plot:
- X (the variable for the x axis)
- Y (the variable for the y axis)
- Color (the variable that determines the color of the points)

## Procedure

1. Chose which variable to put in for X
2. Chose which variable to put in for y 
3. Chose which variable to put in for the color

The plot should automatically create itself after you chose each variable

## Example 
This an example plot that can be created by selecting
X = carat Y = Price and Color = color
```{r, echo=FALSE}
library(ggplot2)
data("diamonds")
p <- ggplot(data = diamonds, aes(x = carat, y = price, color= color)) + geom_point()
p <- p + xlab(as.character("Carat")) + ylab(as.character("Price"))
p
```